---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# teal.widgets R Package Development Guide

## Package Overview

`teal.widgets` is part of the `teal` framework. It provides a set of reusable `shiny`
widgets that `teal` modules use to build their interface: plot and table outputs with
resizing and download buttons, layout helpers, and a few custom inputs. Each widget
works on its own; there is no shared app, router, or state.

The widgets come in two forms:

- **Full modules** are a `<name>_ui()` and `<name>_srv()` pair, for example
  `plot_with_settings`, `table_with_settings`, and `verbatim_popup`.
- **Plain helpers** are single functions, such as `standard_layout`,
  `optionalSelectInput`, `draggable_buckets`, `panel_group`, `white_small_well`, and
  `get_dt_rows`.

## Development Context

Most of the work in this package is about rendering plots and tables and letting the
user resize and download them.

### Relationships with other packages

`teal.widgets` sits at the bottom of the framework. `teal` and the `teal.modules.*`
packages depend on it, but it does not depend on any of them.

Packages it relies on:

- `shiny`, `htmltools`, and `bslib` for the UI and layout.
- `shinyWidgets` and `shinyjs` for richer inputs and client-side behavior.
- `rtables`, `gt`, and `gtsummary` for the table types it renders and exports.
- `ggplot2` and base/grid graphics for plots.
- `rvest` and `xml2` for reading and editing table HTML.
- `checkmate` for input validation.
- `webshot2` (a `Suggests`) is needed to download `gt`, `gtsummary`, or `tbl_split`
  tables as PDF. When it is missing the module warns once per session; set
  `DISABLE_GT_WEBSHOT2_WARNING=true` to turn the warning off.

### How plots and tables are handled

- **Plots.** `plot_with_settings` works out the plot type on its own (`ggplot`,
  lattice, `grob`, or base graphics), so it is not limited to `ggplot2`. Its `plot_r`
  argument can be a `reactive` or a plain `function`; a plain `function` is only needed
  when a base plot has to be captured for download.
- **Tables.** `export_table`, `render_table_to_html`, and `file_download_format` (in
  `R/table_with_settings.R`) use S3 dispatch, with one method per table class
  (`rtables`, `gt`, `gtsummary`, `tbl_split`).
- **Plot and table settings.** `ggplot2_args()` and `basic_table_args()` capture the
  options, `resolve_*()` merges them with the defaults (including the global
  `getOption("teal.ggplot2_args")` and `getOption("teal.basic_table_args")`), and
  `parse_*()` produces the final code.
- **Front-end.** CSS and JS files live in `inst/<widget>/` and are registered with
  `htmltools::htmlDependency()`.

### Workflows

- Before fixing a bug, remember that `teal.widgets` is at the bottom of the framework,
  so the cause is often somewhere else: in the `teal` module that calls the widget, or
  in the package that produced the object being rendered (`rtables`, `gt`, `gtsummary`,
  `ggplot2`). Confirm the problem is really here before changing widget code, and check
  whether an issue already exists.
- To support a new table type, add a method for the new class to each of
  `export_table`, `render_table_to_html`, and `file_download_format`, instead of
  branching on class.
- To change how plot or table settings are applied, edit `ggplot2_args()` /
  `basic_table_args()`, `resolve_*()`, and `parse_*()` together, since a setting flows
  through all three.
- To add front-end assets, put the file in `inst/<widget>/` and register it in the
  widget's dependency function; do not inline a `tags$script`.
- Run the full test suite before trusting a green run: set `TESTING_DEPTH=5`, because
  the default of `3` skips the heavy `shinytest2` tests. Server and unit tests are in
  `test-<widget>.R`; UI tests are in `test-<widget>_ui.R`.
- Add a regression test when you fix a bug.

This package is part of the teal framework. The following configuration applies to all packages within the teal framework:

```{r teal_content, echo=FALSE, results='asis'}
# Append common shared instructions from teal framework
common_claude_url <- "https://raw.githubusercontent.com/insightsengineering/teal/refs/heads/add_common_claude%40main/SHARED_CLAUDE.md" # nolint: line_length_linter.
common_claude_path <- tempfile()
download.file(url = common_claude_url, destfile = common_claude_path, quiet = TRUE)
common_claude_shared <- readLines(common_claude_path)
# remove CI gitlab content and github section comments
gitlab_starts <- grep("gitlab starts here", common_claude_shared)
gitlab_ends <- grep("gitlab ends here", common_claude_shared)
common_claude_shared <- common_claude_shared[-1 * seq(gitlab_starts, gitlab_ends)]
github_starts <- grep("github starts here", common_claude_shared)
github_ends <- grep("github ends here", common_claude_shared)
common_claude_shared <- common_claude_shared[-1 * c(github_starts, github_ends)]
# collapse consecutive blank lines so the rendered markdown passes markdownlint (MD012)
is_blank <- common_claude_shared == ""
common_claude_shared <- common_claude_shared[!(is_blank & c(FALSE, head(is_blank, -1)))]
cat(common_claude_shared, sep = "\n")
```
