---
title: "Custom basic_table arguments module"
author: "Maciej Nasiński"
output: rmarkdown::html_vignette
runtime: shiny
vignette: >
  %\VignetteIndexEntry{Custom basic_table arguments module}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette will guide you through implementation of custom `rtables::basic_table` arguments for `rtables` tables 
based modules. We will enable 2 ways of updating `rtables::basic_table` by the end users.
The `rtables::basic_table` specification could be updated with the `teal.basic_table_args` `options` variable or a 
`ggplot2_args` argument in a `tm_g_*` module. We still take into account default specification set up by the module 
creator in the server function, which has the lowest priority.

The implementation should consist of 5 steps:

1. Adding `basic_table_args` argument to the `tm_t_*` function and then its server function.
The default should be set to
the `basic_table_args()` function for a single plot.
and  `list(default = basic_table_args())` multi-table modules.
2. Adding a validation (e.g. `stopifnot` or `checkmate`) of the `basic_table_args` argument to the `tm_*` function.
The validation is more complex for multi-table modules, where the `ggplot2_args` could be a `list`.
The module creator has to provide a list of plots names, which should be validated at this step and added to the `param` 
field in `roxygen2`. For multi-table modules the step 
`if (is_basic_table_args) basic_table_args <- list(default = basic_table_args)` is recommended.
3. Aggregating and reducing all `basic_table_args` sources with `resolve_basic_table_args()`.
4. Usage of the `parse_basic_table_args()` function which will aggregate and reduce all inputs
to one expression.
5. Adding the created expression to the chunk with a table.

The `parse_basic_table_args()` function picks the first non NULL value for each argument, checking in order:

1. `basic_table_args` argument provided by the end user.
For multi-table case, per table (`basic_table_args_table`) and then default (`basic_table_args_default`) setup.
2. Global R variable (`options`), `teal.basic_table_args`.
3. `basic_table_args_developer` which is a developer setup, lowest priority.

## Example - Single-Table Module

```{r echo=TRUE}
options("teal.basic_table_args" = teal.widgets::basic_table_args(title = "ENV_TITLE"))
library(shiny)
library(teal.widgets)
library(magrittr)

basic_table_args <- list(
  default = basic_table_args(prov_footer = "USER_FOOTER"),
  table1 = basic_table_args(subtitles = "USER_SUBTITLES_TABLE1"),
  table2 = basic_table_args(subtitles = "USER_SUBTITLES_TABLE2")
)

app <- shinyApp(
  ui = fluidPage(
    shinyjs::useShinyjs(),
    div(verbatimTextOutput("table1"))
  ),
  server = function(input, output, session) {
    dev_table_args <- teal.widgets::basic_table_args(show_colcounts = TRUE)

    table_expr <- substitute(
      expr = {
        tt <- f_table_expr %>%
          rtables::split_cols_by("Species") %>%
          rtables::analyze(vars = "Sepal.Length", afun = function(x) {
            rtables::in_rows(
              "Mean" = rtables::rcell(mean(x), format = "xx.xx"),
              "Range" = rtables::rcell(range(x), format = "xx.xx - xx.xx")
            )
          })
        table2 <- rtables::build_table(tt, iris)
        table2
      },
      env = list(f_table_expr = parse_basic_table_args(
        teal.widgets::resolve_basic_table_args(
          user_table = basic_table_args$table2,
          user_default = basic_table_args$default,
          module_table = dev_table_args
        )
      ))
    )
    output$table1 <- renderPrint(eval(table_expr))
  }
)
```

```{r echo=TRUE, eval = FALSE}
shinyApp(app$ui, app$server)
```
